### 2022-08-29 paper/ms.tex

% region preamble

\documentclass[fleqn,usenatbib]{mnras}

\usepackage[T1]{fontenc}

% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{newtxtext,newtxmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
\newcommand{\Vc}{\ensuremath{V_{\textrm{c}} }}
\newcommand{\PS}{\ensuremath{\Omega_{\textrm{p}}}}
\newcommand{\Rb}{\ensuremath{R_{\textrm{b}}}}

\newcommand{\Nbody}{$N$-body}
\newcommand{\AREPO}{\textsc{AREPO}}
\newcommand{\SMUGGLE}{SMUGGLE}
\newcommand{\Msun}{\ensuremath{M_{\odot}}}

\DeclareMathOperator\arcsinh{arcsinh}

% endregion

% region title
%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Stellar Bars and Gas]{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}

\author[A. Beane et al.]{Angus Beane,$^{1}$\thanks{E-mail: angus.beane@cfa.harvard.edu (AB)}
Lars Hernquist,$^{1}$
Elena~D'Onghia,$^{2,3}$
Federico~Marinacci,$^{4}$
Charlie Conroy,$^{1}$
Jia~Qi,$^{5}$\newauthor
Laura~V.~Sales,$^{6}$
Paul~Torrey,$^{5}$
Mark~Vogelsberger$^{7}$
\\
% List of institutions
$^{1}$Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA\\
$^{2}$Department of Physics, University of Wisconsin-Madison, Madison, WI, USA\\
$^{3}$Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA\\
$^{4}$Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy\\
$^{5}$Department of Astronomy, University of Florida, Gainesville, FL, USA\\
$^{6}$Department of Physics \& Astronomy, University of California, Riverside, CA, USA\\
$^{7}$Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2022}

% endregion

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Elongated bar-like features are ubiquitous in galaxies, occurring at the centers
of approximately two-thirds of spiral disks.  Due to gravitational interactions
between the bar and the other components of galaxies, it is expected that
angular momentum and matter will redistribute over long (Gyr) timescales in
barred galaxies. Previous work ignoring the gas phase of galaxies has
conclusively demonstrated that bars should slow their rotation over time due to
their interaction with dark matter halos. We have performed a simulation of a
Milky Way-like galactic disk hosting a strong bar which includes a
state-of-the-art model of the interstellar medium and a live dark matter halo.
In this simulation the bar pattern does not slow down over time, and instead
remains at a stable, constant rate of rotation. This behavior has been observed
in previous simulations using more simplified models for the interstellar gas
but it has remained unexplained. We propose that the gas phase of the disk and
the dark matter halo act in concert to stabilize the bar pattern speed and
prevent the bar from slowing down or speeding up. We find that in a Milky
Way-like disk, a gas fraction of only about 5\% is necessary for this mechanism
to operate. Our result naturally explains why nearly all observed bars rotate
rapidly and is especially relevant for our understanding of how the Milky Way
arrived at its present state.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
galaxies: bar - galaxies: disc - galaxies: evolution - galaxies: kinematics and
dynamics
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize a stellar disk against bar formtaion
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

It is more difficult to reconcile numerical simulations with the observed
pattern speeds of extragalactic bars. Currently, the best technique for
measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This approach has
recently been applied to samples of galaxies from the MaNGA survey
\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These studies confirm
what was found in earlier works, that nearly all extragalactic bars are fast
rotators (i.e., they rotate close to their maximum rotation rate)

This is a problem for theoretical simulations, for which there is ample evidence
that galactic bars should resonantly interact with the dark matter halo, causing
the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
interaction can be understood as an angular form of dynamical friction between
the bar and the dark matter halo. While studied in detail for the bar
\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
still useful review of bar dynamics, see \citet{1993RPPh...56..173S}

Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
radius of corotation \RCR{} is defined for circular orbits as the radius at
which the orbital frequency is equal to the pattern speed, \PS{}, of a given
non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
speeds of bars indicate that nearly all galaxies have $1 < \Rot <
1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
2020MNRAS.491.3655G}.

While the interaction between the bar and a dark matter halo is well-understood
theoretically, this is not the case for the interaction between a bar and a
gaseous disk. Some argue that the gas disk should slow down the bar more
\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
drive gas inwards means the bar should speed up due to the effect of the gas
disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
typically contributes only about $10-20\%$ of the mass of a galaxy at the
present day, one might naively expect it to have a subdominant effect on the
bar. However, because gas is collisional, it can participate in non-resonant
angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
numerical work has shown that the gas phase can have a stronger influence on a
bar than its contribution to the mass of a galaxy would suggest
\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.

We have performed a simulation of a disk galaxy using the finite-volume,
gravito-hydrodynamics code AREPO \citep{2010MNRAS.401..791S}. We use the galaxy
formation model Stars and MUltiphase Gas in GaLaxiEs
\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
evolution in the bar pattern speed over several Gyr when the gas phase is
accounted for and properly modeled. This behavior has been observed in a few
previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
2014MNRAS.438L..81A}. However, we are the first to demonstrate that the behavior
ought to be present in a large fraction of galaxies, since it occurs in our
simulation with gas fractions as low as about $5\%$. We are also the first to
provide a physical explanation of the behavior with clear observational
predictions.

We show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in
a case with and without gas in Fig.~\ref{fig:overview}. We see that the bar
grows longer and stronger without gas, while it remains at approximately the
same length and strength when gas is included. Details on the production of
these synthetic images are given in Appendix~\ref{app:hst}.

% In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
% galaxy exhibits some similarities to the Milky Way, described in more detail in
% Section XX. Surprisingly, we find that the pattern speed of the bar in this
% galaxy is constant with time. We find that with even modest gas fractions of
% about 5\%, our isolated galaxy displays no evolution in pattern speed over many
% $\textrm{Gyr}$.

In Section~\ref{sec:methods}, we describe our initial setup, numerical model,
and details on our bar analysis procedures. In Section~\ref{sec:results}, we
summarize the main results from our findings. We discuss these findings at more
length and in the context of previous research in Section~\ref{sec:discussion}
before concluding in Section~\ref{sec:conclusions}.

% The expectation that a dark matter halo acts to slow down a bar is in conflict
% with observational estimates of bar pattern speeds. Bar rotation rates are
% typically classified by the dimensionless ratio
% % \begin{equation}
% % \Rot = \RCR/R_b\textrm{,}
% % \end{equation}
% where \RCR{} is the radius of corotation and $R_b$ is the length of the
% bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits as
% the radius at which the orbital frequency is equal to the pattern speed,
% $\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
% circular velocity $V_c$, it is given by $\RCR = V_c / \Omega_p$.} Galaxies with
% $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot > 1.4$
% are considered ``slow rotators''\cite{2000ApJ...543..704D}. Galaxies with $\Rot
% < 1$ are not thought to be stable\cite{1980AA....81..198C}. Observational
% estimates of the pattern speeds of bars indicate that nearly all galaxies have
% $1 < \Rot < 1.4$\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
% 2019MNRAS.482.1733G, 2020MNRAS.491.3655G}.

% The role of gas on the evolution of the bar is less well-understood. Since the
% gas phase typically only contributes about $10-20\%$ of the mass of a galaxy at
% the present day, one might naively expect it to have a subdominant effect on the
% bar. However, because gas is collisional, it can participate in non-resonant
% angular momentum exchange with the bar\cite{2011MNRAS.415.1027H}. Thus,
% numerical work has shown that the gas phase can have a stronger influence on a
% bar than its contribution to the mass of a galaxy would
% suggest\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.

\section{Methods}
\label{sec:methods}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
    \caption{The \textit{upper panel} shows the evolution of the pattern speed.
    As expected, the bar in the \Nbody{} run slows down due to interactions
    between the bar and the dark matter halo. However, the bar in the SMUGGLE
    run does not slow down and instead remains at a constant pattern speed. The
    \textit{lower panel} shows the torque on the bar by different components.
    The solid lines indicate the torque exerted by the halo in both the \Nbody{}
    and SMUGGLE cases. The dashed line is the torque exerted by the gas phase in
    the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
    properties are calculated is given in the Methods section.}\label{fig:prop}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/Rb_A2.pdf}
    \caption{The evolution in bar length and strength. The \textit{upper panel}
    shows the evolution of the bar length. In the \Nbody{} case, the bar
    lengthens. This occurs because as the pattern speed drops, bar-like orbits
    at larger radii are possible. Stars are captured on these orbits,
    lengthening the bar. This process does not occur in the SMUGGLE cases since
    the bar pattern speed is not decreasing, and therefore the bar length
    remains constant. The bar strength, shown in the \textit{lower panel} is
    measured as the maximum of the second Fourier component divided by the
    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
    strength increases with time, consistent with previous results showing that
    the bar strength increases as bars slow down. In the SMUGGLE case (orange),
    we see that the bar strength remains roughly constant, possibly slightly
    decreasing with time. This is also consistent with the expected relation
    between pattern speed and strength since the bar in this case is not slowing
    down.}
    \label{fig:strength}
\end{figure}
\subsection{Initial Conditions}
The initial setup of the galactic disk used in this work follows closely the
GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
three components - a radially exponential and vertically isothermal stellar
disk, and a stellar bulge and dark matter halo following a Hernquist profile
\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
parameters as the GALAKOS disk, more details of which can be found in the
original paper.

The addition of the gas phase was done as follows. The version of
\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas disk
which is radially exponential and in vertical gravito-hydrodynamic balance. We
modified the radial profile of this code in order to allow us to generate a disk
with a constant surface density within some cut-off radius, and then
exponentially declining beyond that radius with the scale-length of the stellar
disk. Our fiducial model used an initial surface density of
$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and solar
metallicity.

After generating the gaseous disk in this way, we stitched the gas disk together
with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after the
GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of
allowing the GALAKOS disk to evolve first for a short period of time is to allow
for the bar to form unimpacted by the presence of the gas. We found that
including the gas before the bar has formed disrupts the formation of the bar,
as has been seen previously \citep[e.g.,][]{2013MNRAS.429.1949A}. Throughout
this work, we consider $t=0$ for the \Nbody{} run to be the time at which we
added the gas phase for the SMUGGLE run (i.e., we ignore the first
$1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).

We made one additional modification when stitching the gas disk together with
the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$ of
the gas disks. This hole guards against an initial dramatic infall of gas within
the bar region, which we found to destroy the bar. It is not uncommon for
observed barred galaxies to have gas deficits in the bar region \citep[though
not in the very center;][]{1993RPPh...56..173S}. Therefore, our practice of
allowing the gas distribution to have a hole in the central region is consistent
with our choice to begin the simulations with a bar already formed. In this
manner, we are able to study the ensuing self-consistent interaction between the
bar and the gas, but of course we are unable to explore the origin of bars in
the presence of the gas.

We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
components (initial stellar disk, stellar bulge, and gas) and a mass resolution
of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass resolution is
closest to ``level 3'' in the AURIGA simulations \citep{2017MNRAS.467..179G}.
This corresponds to approximately $6.4\times10^6$ particles in the stellar disk,
$1.1\times10^6$ in the bulge, $1.2\times10^6$ in the gas disk, and
$25.3\times10^6$ in the dark matter halo. We used a softening length of
$0.02\,\textrm{kpc}$ for all components. Snapshots were saved at equal intervals
of $0.005$ in the time units of the simulation,
$\textrm{kpc}/(\textrm{km}/\textrm{s})$.

Our setup is initially out of equilibrium, but we found that after about
$500\,\textrm{Myr}$, the system has settled into a roughly steady-state
configuration and initial transients appear not to affect the results after this
point. The constant surface density of the initial gas disk is important for
ensuring that the gas disk is dense enough in order for comparisons to real
galaxies to be appropriate.

\subsection{Numerical Model}
We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
\citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
model additionally includes radiative heating and cooling, star formation, and
stellar feedback. Explicit gas cooling and heating of the multi-phase
interstellar medium is implemented, covering temperature ranges between $10$ and
$10^8\,\textrm{K}$.

Star formation occurs in cells above a density threshold
($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
$\epsilon = 0.01$. Star formation converts gas cells into star particles which
represent single stellar populations with a Chabrier initial mass function
\citep{2003PASP..115..763C}. For each star particle, the deposition of energy,
momentum, mass, and metals from stellar winds and supernovae is modeled.
Photo-ionization and radiation pressure are handled using an approximate
treatment. A more detailed description of this model can be found in the
flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.

We used the fiducial model parameters, except that we increased the number of
effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from $64$
to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
inefficient photo-ionization feedback since the photo-ionizing budget had not
been exhausted after deposition into $64$ neighboring cells. We also employed an
updated version of SMUGGLE using a new mechanical feedback routine similar to
the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
tensor renormalization which ensures linear and angular momentum conservation to
machine precision.

In addition to the SMUGGLE model, we considered a simpler model of the
interstellar medium based upon \citet{2003MNRAS.339..289S}. In this approach,
the multiphase nature of the interstellar medium is described in a subgrid
manner by allowing each resolution element to have a ``cold'' and ``hot''
component, with the equation of state of the gas suitably modified. Gas is
allowed to interchange between the cold and hot components through processes
such as cooling and stellar feedback. Cold gas is allowed to undergo star
formation. We refer to this model as the smooth interstellar medium model, and
it is described in more detail in \citet{2019MNRAS.489.4233M}.

\subsection{Bar Analysis}
The analysis of various bar properties is performed as follows. First, the
pattern speed is measured from the angle of the second Fourier component. We
measured the second Fourier component by computing,
\begin{equation}
\begin{split}
A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
A_0 &= \sum_i m_i \textrm{,}
\end{split}
\end{equation}
where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
After correcting for the periodicity of $\phi_b$, we measured the pattern
speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
time.

In order to compute other properties of the bar, it is necessary to decompose
the disk into a barred and unbarred component. We achieved this by following
closely the methods described in \citet{2016MNRAS.463.1952P}. Our implementation
is described in more detail in Appendix~\ref{app:bardecomp}. After the disk has
been decomposed into a trapped and untrapped component, we measured the bar
length as being the radius $R_b$ which encapsulates $99\%$ of the stars
identified as being trapped in the bar.

To compute torques we used the tree algorithm in \texttt{MakeNewDisk}
\citep{2005MNRAS.361..776S} customized to be accessible from \texttt{Python}
using \texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code
\citep{1989ApJS...70..419H}. We constructed a tree with an opening angle of
$0.35$ using only the star particles identified as being trapped in the bar. We
then queried the tree at the locations of all resolution elements in the other
components and computed the torque of the bar on such components. The torque on
the bar by the other components is simply the negative of the torque on the
other components by the bar.

\subsection{Plotting Details}
We saved snapshots in intervals of $0.005$ in the time units of the
simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
throughout this work we referred to the native code time unit as
$\textrm{Gyr}$. None of our results are sensitive to this choice.

In order to remove numerical noise in several quantities we computed, we
applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
filter was applied to plots of pattern speeds, bar lengths and strengths,
torques, and angle differences.

\section{Results}
\label{sec:results}

We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
while the pattern speed in the SMUGGLE case remains roughly constant. The
slowing down of the pattern speed in the \Nbody{} case is consistent with a long
line of numerical research on bars in \Nbody{} simulations
\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
2009ApJ...697..293D}.

However, in the SMUGGLE case the pattern speed remains constant. After the first
Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
pattern speed for the \Nbody{} run over the same interval.

The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
different components. The solid lines indicate the torque on the bar by the dark
matter halo, whereas the dashed line is the torque on the bar by the gas phase.
In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
exerts a similar negative torque on the bar in the first Gyr of evolution, but
after that the halo exerts a much weaker torque on the bar, averaging only
$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
the same units.

As we saw qualitatively in Fig.~\ref{fig:overview}, the middle panel of
Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
grows over time while it remains roughly constant in the SMUGGLE case. This is
also consistent with previous numerical work, which found that bars tend to grow
as they slow down and the radius of corotation increases
\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
radius, is shown in the lower panel Fig.~\ref{fig:strength}. The quantity
$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
increases over time as the bar pattern slows. This is consistent with previous
\Nbody{} simulations which showed a clear correlation between the bar pattern
speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
case, we see that the bar strength has an initial drop but then remains at a
roughly constant, but slightly decreasing, strength. This is consistent with the
pattern speed in the SMUGGLE case being roughly constant or slightly increasing.

\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/halo_wake.pdf}
    \caption{The wake excited in the dark matter halo. The dark matter halo wake
    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
    \textit{middle} and \textit{right panels} show a surface density projection
    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
    has been subtracted. The solid line indicates the direction of the bar while
    the dashed line indicates the direction of the halo wake (both measured by
    taking the second Fourier component within a sphere of all material within a
    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
    evolution of the angle difference between the bar and the halo wake, as
    measured from the second Fourier component. After the first Gyr, the angle
    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
    \label{fig:wake}
\end{figure*}

% \subsection{Figures and tables}

% Figures and tables should be placed at logical positions in the text. Don't
% worry about the exact layout, which will be handled by the publishers.

% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
% e.g. Table~\ref{tab:example_table}.

% % Example figure
% \begin{figure}
% 	% To include a figure from a file named example.*
% 	% Allowable file formats are eps or ps if compiling using latex
% 	% or pdf, png, jpg if compiling using pdflatex
% 	\includegraphics[width=\columnwidth]{example}
%     \caption{This is an example figure. Captions appear below each figure.
% 	Give enough detail for the reader to understand what they're looking at,
% 	but leave detailed discussion to the main body of the text.}
%     \label{fig:example_figure}
% \end{figure}

% % Example table
% \begin{table}
% 	\centering
% 	\caption{This is an example table. Captions appear above each table.
% 	Remember to define the quantities, symbols and units used.}
% 	\label{tab:example_table}
% 	\begin{tabular}{lccr} % four columns, alignment for each
% 		\hline
% 		A & B & C & D\\
% 		\hline
% 		1 & 2 & 3 & 4\\
% 		2 & 4 & 6 & 8\\
% 		3 & 5 & 7 & 9\\
% 		\hline
% 	\end{tabular}
% \end{table}

\section{Discussion}
\label{sec:discussion}
\subsection{Pattern Speed Evolution}
The lack of evolution in the pattern speed of the SMUGGLE case (seen in
Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
on the bar by the dark matter halo. We argue that this can be understood in
terms of the halo wake mechanism. In the \Nbody{} case, halo material that is
resonant with the bar will form a wake which lags behind and exerts a negative
torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
body, it is not guaranteed that a negative torque will slow it down - e.g. a
negative torque could shred the bar, reducing its moment of inertia without
changing its pattern speed. However, the bar seems to empirically respond to a
negative torque induced by a halo wake by slowing down.} As the bar slows down,
the location of the resonances in phase space changes, allowing halo material
newly resonant with the bar to participate in forming the wake. However, the gas
is a reliable source of positive torque on the bar, speeding the bar up. This
stops the resonance location from changing such that the halo cannot reinforce
the wake, arresting the process by which the halo can slow the bar down. We term
this process ``exhaustion.''

We can test this interpretation by measuring the angle offset between the halo
wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
offset), then the wake cannot exert a negative torque on the bar. This angle
is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
wake with respect to the location of the bar in the \Nbody{} (center) and SMUGGLE
(right) cases at one point in time.

The presence of the gas can arrest the process by which additional material in
the dark matter halo can contribute to a wake. However, this does not explain
why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
Naively, it would be a coincidence that the bar pattern speed remains constant
in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
torques. However, a constant pattern speed in the presence of gas has been
observed in a few simulations of barred galaxies with
gas \citep{1993AA...268...65F, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. Previous
work has argued this is due to the bar torquing gas inwards, but no explanation
has been given for why it might remain constant.

We propose that an equilibrium mechanism is responsible for the pattern speed
remaining approximately constant. In this scenario, residual negative torque
from the dark matter halo balances out the positive torque from the gas phase.
It has been shown when an analytic bar is forced to rotate at a constant pattern
speed for a few Gyr, the halo exerts almost no torque on the bar
\citep{2022MNRAS.513..768C}. We saw in Fig.~\ref{fig:prop} that the dark matter
halo in our simulation is still able to support some negative torque over a
several Gyr time span.

We argue that the following occurs. First, the bar is not able to slow down
quickly enough due to the positive torque of the infalling gas. This causes
the resonant halo phase space at a particular pattern speed, $\Omega_{p,0}$,
to become exhausted and no longer able to support a negative torque. Second,
the gas is still exerting a positive torque on the bar, and therefore the
pattern speed will again increase. Since at higher pattern speeds the halo has
not yet been totally exhausted, the halo will once again be able to exert a
negative torque on the bar. The pattern speed will then settle at a new value
slightly higher than $\Omega_{p,0}$ where the gas and halo torques cancel.
Over time, the pattern speed should slowly increase as the halo becomes
progressively exhausted.

\subsection{Delayed Gas Injection}
A clear prediction of our proposed mechanism is that the constant pattern speed
a particular galaxy will end up is somewhat arbitrary. In the real universe for
an isolated galaxy it would be the formation pattern speed of the bar while in
our simulation it is the pattern speed of the bar when gas is added to the
system. We tested this by adding gas to the system at a later time when the bar
has further grown and slowed down with time. In our particular test, we added
the gas at a time when the pattern speed is
$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very similar
between the two cases (orange and red lines). If anything, the system with a
lower pattern speed seems to speed up more, which is consistent with our picture
since the stronger bar should experience a larger torque from the gas as it is
more efficient at driving gas inflows. We also show in the
Appendix~\ref{app:varyps} that more slowly rotating bars at fixed bar strength
are more efficient at driving gas inflows as well. Nonetheless, when the initial
pattern speed is lower (red line), the addition of gas does not cause the
pattern speed to quickly return to the higher value of our fiducial simulation
(orange line).

\subsection{Varying Initial Gas Fractions}
We also performed a test in which we varied the initial gas fraction of the
disk. In our fiducial run, we set the surface density of the gas disk from
$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{pc}^2$.
We also ran with surface densities of $15$, $10$, and
$5\,\Msun/\textrm{pc}^2$. These correspond to initial gas fractions of
approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
shown in Fig.~\ref{fig:fgas} We find that the bar in disks with
initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
with a constant pattern speed while the bar in a disk with initial surface
density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.

As a result, we conclude that for the disk, bar, and halo properties
considered in this work, a gas fraction of only approximately $5\%$ is
necessary in order for the proposed stabilizing mechanism to operate.

\subsection{Smooth Interstellar Medium}
We also performed a simulation of the same disk but with a simpler model of the
interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
in cosmological simulations of galaxy formation and described in more detail in
Section~\ref{sec:methods}. The result of this test is presented in
Fig.~\ref{fig:GFM}. We find that the pattern speed evolution is nearly the same
in this case, and so conclude that our result is not sensitive to the details of
the model for the interstellar medium.

\subsection{Observations}
Observational estimates of the pattern speeds of bars indicate that nearly all
galaxies have $1 < \Rot < 1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A,
2019MNRAS.482.1733G, 2020MNRAS.491.3655G}, where $\Rot$ was defined in
Section~\ref{sec:intro} to be $\Rot\equiv \RCR/\Rb$. This observational fact has
long been in conflict with the theoretical expectation that bars should slow
down, increasing \Rot{} \citep[e.g.][]{1984MNRAS.209..729T, 1985MNRAS.213..451W,
2000ApJ...543..704D}. Explanations for this discrepancy have been given in the
past. Some have argued that perhaps the central regions of dark matter haloes
are less dense than we expected from $\Lambda\textrm{CDM}$
\citep[e.g.][]{2000ApJ...543..704D,2021AA...650L..16F}. Some have argued that
perhaps bars are recurrent, short-lived phenomena, and that all the bars we see
in the local universe are very young \citep{2002AA...392...83B,
2005MNRAS.364L..18B}. Some have argued that modifications to General Relativity
ease the tension between the observed universe and $\Lambda\textrm{CDM}$
\citep[e.g.][]{2021MNRAS.503.2833R, 2021MNRAS.508..926R}.

Because such a small gas fraction is necessary for our stabilizing mechanism to
operate ($5\%$ in our Milky Way-like disk), we argue that most galaxies host a
bar that is not slowing down. This naturally explains why most observed bars are
fast rotators. However, we acknowledge two instances of reported discrepancies
between our mechanism and observations.

First, we note that \citet{2020MNRAS.491.3655G} found that the rotation
parameter \Rot{} positively correlates with gas fraction, such that galaxies
with higher gas fractions are rotating more slowly. However, it is not obvious
this is in tension with our result since the gas fraction of galaxies correlates
with other galactic properties \citep{2009ARAA..47..159B}. Furthermore, the
measurement of pattern speeds is a delicate process still prone to large errors.
\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
    \caption{Pattern speed evolution with a lower initial pattern speed. We
    tested the evolution of our system when gas is added to the \Nbody{} run at
    a later time, but with all other simulation parameters kept the same. The
    setup is therefore identical to our previous runs just with a bar that is
    larger, stronger, and with a lower pattern speed. We find that the pattern
    speed evolution is very similar to our fiducial case, except that the bar
    retains its original pattern speed. This indicates a mechanism which keeps
    the bar at its formation pattern speed, and that there is not a particular
    pattern speed which the system tends to.}
    \label{fig:snap700}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
    \caption{Pattern speed evolution with varying gas fractions. We explored the
    impact of lowering the initial gas surface density of our fiducial disk on
    the evolution of the pattern speed. The surface densities we tested of $20$,
    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
    remain at a constant pattern speed, while an initial surface density of
    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
    same manner as the \Nbody{} case.}
    \label{fig:fgas}
\end{figure}

Second, the work of \citet{2021MNRAS.500.4710C} and \citet{2021MNRAS.505.2412C}
have made indirect measurements of the deceleration of the bar's pattern speed
from kinematics and chemistry. We point out that these reported measurements are
not direct measurements of the Milky Way bar's deceleration. Much like the
simulations in the present work, the simulations of these two works do not
properly account for the complicated formation process of the Galactic bar,
which may leave imprints on the present day distribution of stars in spatial,
kinematic, and chemical space. More investigation is necessary to reconcile the
present work with these two well-executed manuscripts.

\subsection{Cosmological Simulations}
Barred galaxies in cosmological simulations of galaxy formation continue to be
in conflict with observations by producing bars which rotate too slowly
\citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}. However,
the pattern speeds of bars in both cosmological simulations and the real
universe can be affected by environmental processes not included in our
simulation -- e.g., satellite infall \citep{2011Natur.477..301P}, non-sphericity
\citep{2013MNRAS.429.1949A}, rotation in the dark matter halo
\citep{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
2019MNRAS.488.5788C}, or perhaps even the gaseous circumgalactic medium.
Naturally, extending our present work to account for such effects is a crucial
next step in understanding the formation and evolution of galactic bars.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
    which treats its multiphase nature in a subgrid fashion
    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
    model, which explicitly resolves the hot and cold phases of the ISM
    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
    broadly similar to the evolution in the SMUGGLE case. This shows that the
    stability of the pattern speed is not simply a result of our assumed model
    for the ISM.}
\label{fig:GFM}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense \citep{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics \citep{2021MNRAS.503.2833R,
2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar \citep{2021MNRAS.500.4710C,
2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time Our work does not alter
expectations for radial mixing induced by spiral arms
\citep{2002MNRAS.336..785S}.

We found that below a certain gas fraction, bars should still be able to slow
down. Therefore, we expect barred spiral galaxies which have been gas-poor for
extended periods of time to be rotating very slowly. We therefore predict that
observations which target such galaxies (e.g., lenticular barred galaxies
\citep{2009ARAA..47..159B}) would find slowly rotating bars.\footnote{This does
not mean that we predict \textit{all} gas-poor galaxies should be slowly
rotating. Indeed, they would need to be gas-deficient for several $\textrm{Gyr}$
before they would be classified as slow rotators.} There does exist one example
of a galaxy which is known to be a slow rotator -- the low surface brightness
galaxy UGC 628 \citep{2009AA...499L..25C}. This galaxy has been studied in
detail by \citet{2016MNRAS.463.1751C}, who note that it indeed has a low gas
fraction for galaxies of its type. We predict a general trend that bars in
gas-rich spiral galaxies should rotate quickly while some bars in gas-poor
spiral galaxies should rotate slowly.

\section*{Acknowledgements}
We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Joshua~S. Speagle,
Martin~D. Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to
thank Todd Phillips for helpful discussions. Some computations in this paper
were run on the FASRC Cannon cluster supported by the FAS Division of Science
Research Computing Group at Harvard University. Resources supporting this work
were also provided by the NASA High-End Computing (HEC) Program through the NASA
Advanced Supercomputing (NAS) Division at Ames Research Center. This research
has made use of the Spanish Virtual Observatory (https://svo.cab.inta-csic.es)
project funded by MCIN/AEI/10.13039/501100011033/ through grant
PID2020-112949GB-I00. A.B. was supported by the Future Investigators in NASA
Earth and Space Science and Technology (FINESST) award number 80NSSC20K1536
during the completion of this work. E.D. was partially supported by HST grants
HST-AR-16363.001 and HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761.
J.Q. acknowledges support from NSF grant AST-2008490. L.V.S. is grateful for
financial support from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER
1945310 grants. P.T. acknowledges support from NSF grant AST-1909933,
AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges support through
NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants
AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Data Availability}

The inclusion of a Data Availability Statement is a requirement for articles published in MNRAS. Data Availability Statements provide a standardised format for readers to understand the availability of data underlying the research results described in the article. The statement may refer to original data generated in the course of the study or to third-party data analysed in the article. The statement should describe and provide means of access, where possible, by linking to the data or providing the required accession numbers for the relevant databases or DOIs.

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{ref} % if your bibtex file is called example.bib

% Alternatively you could enter them by hand, like this:
% This method is tedious and prone to error if you have lots of references
%\begin{thebibliography}{99}
%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
%Others S., 2012, Journal of Interesting Stuff, 17, 198
%\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Bar Decomposition}
\label{app:bardecomp}
\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/bar_decomp.pdf}
    \caption{Disk decomposition into the barred and unbarred disk. This
    procedure is based on \citet{2016MNRAS.463.1952P}. The \textit{left panel}
    shows a face-on surface density projection through the stellar component of
    the \SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
    \textit{middle panel} shows the component of the disk identified as being
    trapped in the bar while the \textit{right panel} shows the component of the
    disk identified as not being trapped in the bar. The fact that the untrapped
    stars form a roughly axisymmetric structure indicates our bar decomposition
    is sufficiently accurate.}
    \label{fig:decomp}
\end{figure*}

Computing the length of the bar and the torque on the bar by different
components requires us to decompose the disk into a component which is trapped
by the bar and a component which is untrapped. In order to do this, we follow
closely the technique developed in \citet{2016MNRAS.463.1952P}. We analyzed the
orbit of each star particle (meaning initial disk, bulge, and newly formed
stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
corotating with the bar, where apoapses are defined as local maxima in $r$. For
each apoapse, we searched for the $19$ closest apoapses in time and applied a
$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
computed for each of the two clusters the average angle from the bar
$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
${\sigma_R}_{0,1}$, and the average radius of the cluster
$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in the
bar if it met the following criteria:
\begin{equation}
\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
\end{equation}
\begin{equation}
\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
\end{equation}
These criterion are slightly different and simplified from the ones used in \citet{2016MNRAS.463.1952P}, but we found them to empirically work well at
decomposing the disk into a bar and disk component. In Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
\textit{left} panel shows a surface density projection of the stellar disk and
bulge (including newly formed stars) from the SMUGGLE model after
$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
that are identified as being trapped in the bar and the \textit{right} panel
shows a projection of the stars that are not identified as being trapped. The
fact that the \textit{right} panel is roughly axisymmetric indicates the bar
decomposition is performing adequately.

\section{Varying Pattern Speed}
\label{app:varyps}
When the bar slows down, we argue that this induces a larger positive torque
from the gas phase. Only gas within corotation will flow inwards, while gas
outside corotation will flow outwards \citep{2011MNRAS.415.1027H}. Since the
corotation radius is larger for more slowly rotating bars, it follows that
more slowly rotating bars should be more efficient at driving gas inflows and
thus experience a larger positive torque from the gas phase.

We performed an experiment to test this hypothesis by forcing the stellar disk
in the SMUGGLE run to rotate at a constant angular rate and measuring the
torque on the bar by the gas phase at different rotation rates. The result of
this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
slowly rotating bar experiences a larger positive torque from the gas.

\begin{figure}[h!]
    \centering
    \includegraphics[width=9cm]{fig/torque_ps.pdf}
    \caption{Average torque exerted by gas on a bar which rotates at a fixed
    pattern speed. Since only gas within the corotation radius is able to infall
    and slower bars have larger corotation radii, slower bars experience a
    larger net torque than faster bars. The setup of the simulations used here
    is identical to the SMUGGLE case discussed earlier, except the \Nbody{} disk
    is rotated as a solid body with a constant angular
    velocity.}
    \label{fig:equil}
\end{figure}

\section{Idealized Mock Light Images}
\label{app:hst}
We created synthetic mock light images at $1024^2$ resolution in
Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
including secondary emission from dust \citep{2020AC....3100381C}. Emission from
star particles was calculated assuming a Bruzual-Charlot spectral energy
distribution \citep{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
function \citep{2003PASP..115..763C}. The luminosity from each star particle was
smoothed using a cubic spline kernel with each star particle being assigned a
smoothing length equal to its distance to its $32$nd closest neighbor. We
launched $10^9$ photon packets per simulation segment. For star particles
which were present in the initial simulation (i.e., initial disk and bulge
particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
metallicity. For star particles in the SMUGGLE simulation which are newly
formed, we used their recorded ages and metallicities.

For the SMUGGLE simulations, we included absorption and emission from dust. We
assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
Solids) dust model \citep{2017AA...602A..46J} to convert gas metallicity and
mass density to dust mass density. For the \Nbody{} simulations we ran the
SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
same procedure.

We used the profiles for filters F814W, F606W, and F475W on the Advanced
Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
scaling following \citet{2004PASP..116..133L} Scalings and cutoff values
were adjusted manually to create visually appealing images. Filter profiles
were provided by the Spanish Virtual Observatory Filter
Service \citep{2012ivoa.rept.1015R, 2020sea..confE.182R}.

\section{Comparison to the Milky Way}
\label{app:milkyway}
For several Gyr, our fiducial disk exhibits several properties in reasonable
agreement to the Milky Way. This is uncommon in models of galaxies that include
the gas phase of the disk but no circumgalactic medium. As mentioned earlier,
the pattern speed seems to match the observed pattern speed of the Milky Way's
bar \citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our
disk is comparable to the Milky Way.

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package \citep{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
bulge, gas, and newly formed stars) with an axisymmetric cylindrical spline with
$20$ grid points in both the radial and vertical direction spanning $0.2$ to
$50\,\textrm{kpc}$ in the radial direction and from $0.02$ to $10\,\textrm{kpc}$
in the vertical direction. We fit the dark matter halo using a spherically
symmetric multipole fit with a maximum angular harmonic coefficient of $l=2$. We
plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
observational estimates \citep{2019ApJ...871..120E}. The SMUGGLE disk (which
includes additional mass in the form of gas) has a slightly higher circular
velocity than the \Nbody{} disk which, itself, is slightly higher than the
observational estimates. Overall, though, the circular velocity curves between
our model and that observed in the Milky Way are broadly consistent.

We also show the evolution of the surface density profile in Fig.~\ref{fig:surf}
We find that in our simulation the atomic and molecular gas surface density and
the SFR surface density is broadly consistent with the expected values for the
Milky Way \citep{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
to the fact that the distances to molecular clouds which underlines this work
used a simple kinematic distance based on an axisymmetric model of the Milky
Way \citep{2017ApJ...834...57M}, which is not accurate in the bar region where gas
has large non-circular velocities.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/surf_dens.pdf}
    \caption{The time evolution of the atomic gas surface density
    (\textit{upper}), molecular gas surface density (\textit{middle}) and the
    star formation rate (SFR) surface density (\textit{lower}) at various times
    during our fiducial simulation. Colored lines indicate the profiles at
    selected times during the simulation while the black dashed lines indicate
    observations for the atomic gas \citep{2008AA...487..951K} and black dotted
    lines indicate a model which allows the CO-to-H$_2$ conversion factor
    $X_{\textrm{CO}}$ to vary with metallicity \citep{2022ApJ...929L..18E}.
    Molecular gas surface densities were provided separately (N. Evans, private
    communication). We see that the molecular gas and SFR surface densities are
    within an order of magnitude of the Milky Way's typical values at all times.
    We see a sharp decrease in the gas and SFR surface densities along the
    extent of the bar from $\sim1$ to $\sim4\,\textrm{kpc}$, related to the gas
    inflow in this region.}
    \label{fig:surf}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/vcirc.pdf}
    \caption{The circular velocity curve of our initial setups. This
    curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
    compared to observational estimates for the Milky
    Way \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
    both runs is marginally larger than the Milky Way's, but still comparable. The
    SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
    additional mass in the gas phase.}
    \label{fig:vcirc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}

\end{document}

% End of mnras_template.tex