### 2022-11-30 paper-ApJ/ms.tex
Add paper-ApJ

%% Beginning of file 'sample631.tex'
%%
%% Modified 2021 March
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.31 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls and other external packages such as
%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
%% AASTeX v6+ still uses v4.1. All of these external packages should 
%% already be present in the modern TeX distributions but not always.
%% For example, revtex4.1 seems to be missing in the linux version of
%% TexLive 2020. One should be able to get all packages from www.ctan.org.
%% In particular, revtex v4.1 can be found at 
%% https://www.ctan.org/pkg/revtex4-1.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%% using aastex version 6.3
\documentclass[twocolumn,linenumbers]{aastex631}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%   tighten      : Makes baselineskip slightly smaller, only works with 
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for 
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much 
%%                  longer author list but may be desirable for short 
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%

\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{newtxtext,newtxmath}
\usepackage{hyperref}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
\newcommand{\Vc}{\ensuremath{V_{\textrm{c}} }}
\newcommand{\PS}{\ensuremath{\Omega_{\textrm{p}}}}
\newcommand{\Rb}{\ensuremath{R_{\textrm{b}}}}
\newcommand{\kms}{\ensuremath{\textrm{km}/\textrm{s}}}

\newcommand{\Nbody}{$N$-body}
\newcommand{\AREPO}{\textsc{AREPO}}
\newcommand{\SMUGGLE}{SMUGGLE}
\newcommand{\Msun}{\ensuremath{M_{\odot}}}
\newcommand{\Myr}{\ensuremath{\textrm{Myr}}}

\DeclareMathOperator\arcsinh{arcsinh}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{March 1, 2021}
%\revised{April 1, 2021}
%\accepted{\today}

\shorttitle{Stellar Bars and Gas}
\shortauthors{Beane et al.}

\graphicspath{{./}{figures/}}

\begin{document}

\title{Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.31 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affiliations. The change was necessary to implement 
%% auto-indexing of affiliations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% argument which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
%% present or in the LaTeX pathway, the OrcID icon will appear next to
%% the authors name.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.31 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.31 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

%\correspondingauthor{August Muench}
%\email{greg.schwarz@aas.org, gus.muench@aas.org}

\author{Angus Beane}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}

\author{Lars Hernquist}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}

\author{Elena D'Onghia}
\affiliation{Department of Physics, University of Wisconsin-Madison, Madison, WI, USA}
\affiliation{Department of Astronomy, University of Wisconsin-Madison, Madison, WI, USA}

\author{Federico Marinacci}
\affiliation{Department of Physics \& Astronomy `Augusto Righi', University of Bologna, Bologna, Italy}

\author{Charlie Conroy}
\affiliation{Center for Astrophysics $|$ Harvard \& Smithsonian,  Cambridge, MA, USA}

\author{Jia Qi}
\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}

\author{Laura V. Sales}
\affiliation{Department of Physics \& Astronomy, University of California, Riverside, CA, USA}

\author{Paul Torrey}
\affiliation{Department of Astronomy, University of Florida, Gainesville, FL, USA}

\author{Mark Vogelsberger}
\affiliation{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, USA}

\begin{abstract}

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral disks.  Due to gravitational
  interactions between the bar and the other components of galaxies, it is
  expected that angular momentum and matter will redistribute over long (Gyr)
  timescales in barred galaxies. Previous work ignoring the gas phase of
  galaxies has conclusively demonstrated that bars should slow their rotation
  over time due to their interaction with dark matter halos. We have performed a
  simulation of a Milky Way-like galactic disk hosting a strong bar which
  includes a state-of-the-art model of the interstellar medium and a live dark
  matter halo. In this simulation the bar pattern does not slow down over time,
  and instead remains at a stable, constant rate of rotation. This behavior has
  been observed in previous simulations using more simplified models for the
  interstellar gas, but the apparent lack of secular evolution has remained
  unexplained. We propose that the gas phase of the disk and the dark matter
  halo act in concert to stabilize the bar pattern speed and prevent the bar
  from slowing down or speeding up. We find that in a Milky Way-like disk, a gas
  fraction of only about 5\% is necessary for this mechanism to operate. Our
  result naturally explains why nearly all observed bars rotate rapidly and is
  especially relevant for our understanding of how the Milky Way arrived at its
  present state.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
%% https://astrothesaurus.org
%% You will be asked to selected these concepts during the submission process
%% but this old "keyword" functionality is maintained in case authors want
%% to include these concepts in their preprints.
\keywords{Classical Novae (251) --- Ultraviolet astronomy(1736) --- History of astronomy(1868) --- Interdisciplinary astronomy(804)}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction}
\label{sec:intro}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

Approximately two-thirds of spiral disks host an elongated bar-like feature at
their centers \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our
own Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of
bars is not difficult to explain, since stellar disks simulated in isolation
almost always form bar-like structures \citep{1971ApJ...168..343H}. Several
studies have shown that a spherical potential, such as a stellar bulge or dark
matter halo, acts to stabilize stellar disks against bar formtaion
\citep[e.g.,][]{1973ApJ...186..467O, 1976AJ.....81...30H}.

It is more difficult to reconcile numerical simulations with the observed
pattern speeds of extragalactic bars. Currently, the best technique for
measuring the pattern speeds of individual galaxies is the Tremaine-Weinberg
method \citep{1984ApJ...282L...5T, 2011MSAIS..18...23C}. This approach has
recently been applied to samples of galaxies from the MaNGA survey
\citep{2019MNRAS.482.1733G, 2020MNRAS.491.3655G}. These studies confirm
what was found in earlier works, that nearly all extragalactic bars are fast
rotators (i.e., they rotate close to their maximum rotation rate).

This is a problem for theoretical simulations, for which there is ample evidence
that galactic bars should resonantly interact with the dark matter halo, causing
the bar to slow down over time \citep{1992ApJ...400...80H, 2000ApJ...543..704D,
2002MNRAS.330...35A, 2002ApJ...569L..83A, 2003MNRAS.341.1179A,
2003MNRAS.346..251O, 2005MNRAS.363..991H, 2006ApJ...637..214M,
2007MNRAS.375..460W, 2009ApJ...697..293D}. The physical mechanism of this
interaction can be understood as an angular form of dynamical friction between
the bar and the dark matter halo. While studied in detail for the bar
\citep{1984MNRAS.209..729T, 1985MNRAS.213..451W}, this process is generic for
any non-axisymmetric disturbance \citep{1972MNRAS.157....1L}. For an old but
still useful review of bar dynamics, see \citet{1993RPPh...56..173S}

Bar pattern speeds are usually measured using the parameter $\Rot=\RCR/\Rb$,
where \RCR{} is the corotation radius and \Rb{} is the bar length.\footnote{The
radius of corotation \RCR{} is defined for circular orbits as the radius at
which the orbital frequency is equal to the pattern speed, \PS{}, of a given
non-axisymmetric feature. In a galaxy with a constant circular velocity \Vc{},
it is given by $\RCR = \Vc / \PS$.} Galaxies with $\Rot < 1.4$ are considered
``fast rotators'' while galaxies with $\Rot > 1.4$ are considered ``slow
rotators'' \citep{2000ApJ...543..704D}. Galaxies with $\Rot < 1$ are not thought
to be stable \citep{1980AA....81..198C}. Observational estimates of the pattern
speeds of bars indicate that nearly all galaxies have $1 < \Rot <
1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G,
2020MNRAS.491.3655G}.

While the interaction between the bar and a dark matter halo is well-understood
theoretically, this is not the case for the interaction between a bar and a
gaseous disk. Some argue that the gas disk should slow down the bar more
\citep{2003MNRAS.341.1179A}, while others argue that the tendency of the bar to
drive gas inwards means the bar should speed up due to the effect of the gas
disk \citep{2013MNRAS.429.1949A, 2014MNRAS.438L..81A}. Since the gas phase
typically contributes only about $10-20\%$ of the mass of a galaxy at the
present day, one might naively expect it to have a subdominant effect on the
bar. However, because gas is collisional, it can participate in non-resonant
angular momentum exchange with the bar \citep{2011MNRAS.415.1027H}. Thus,
numerical work has shown that the gas phase can have a stronger influence on a
bar than its contribution to the mass of a galaxy would suggest
\citep{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.

We have performed a simulation of a disk galaxy using the finite-volume,
gravito-hydrodynamics code \AREPO{} \citep{2010MNRAS.401..791S}. We use the galaxy
formation model Stars and MUltiphase Gas in GaLaxiEs
\citep[SMUGGLE;][]{2019MNRAS.489.4233M}. This disk galaxy exhibits almost no
evolution in the bar pattern speed over several Gyr when the gas phase is
accounted for and properly modeled. This behavior has been observed in a few
previous works \citep{1993AA...268...65F, 2010ApJ...719.1470V,
2014MNRAS.438L..81A}. However, we are the first to demonstrate that the behavior
ought to be present in a large fraction of galaxies, since it occurs in our
simulation with gas fractions as low as about $5\%$. We are also the first to
provide a physical explanation of the behavior with clear observational
predictions.

We show synthetic \textit{Hubble Space Telescope} images of our barred galaxy in
a case with and without gas in Fig.~\ref{fig:overview}. We see that the bar
grows longer and stronger without gas, while it remains at approximately the
same length and strength when gas is included. Details on the production of
these synthetic images are given in Appendix~\ref{app:hst}.

% In this work, we present a simulation of an isolated, gas-rich disk galaxy. Our
% galaxy exhibits some similarities to the Milky Way, described in more detail in
% Section XX. Surprisingly, we find that the pattern speed of the bar in this
% galaxy is constant with time. We find that with even modest gas fractions of
% about 5\%, our isolated galaxy displays no evolution in pattern speed over many
% $\textrm{Gyr}$.

In Section~\ref{sec:methods}, we describe our initial setup, numerical model,
and details on our bar analysis procedures. In Section~\ref{sec:results}, we
summarize the main results from our findings. We discuss these findings at more
length and in the context of previous research in Section~\ref{sec:discussion}
before concluding in Section~\ref{sec:conclusions}.

% The expectation that a dark matter halo acts to slow down a bar is in conflict
% with observational estimates of bar pattern speeds. Bar rotation rates are
% typically classified by the dimensionless ratio
% % \begin{equation}
% % \Rot = \RCR/R_b\textrm{,}
% % \end{equation}
% where \RCR{} is the radius of corotation and $R_b$ is the length of the
% bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits as
% the radius at which the orbital frequency is equal to the pattern speed,
% $\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
% circular velocity $V_c$, it is given by $\RCR = V_c / \Omega_p$.} Galaxies with
% $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot > 1.4$
% are considered ``slow rotators''\cite{2000ApJ...543..704D}. Galaxies with $\Rot
% < 1$ are not thought to be stable\cite{1980AA....81..198C}. Observational
% estimates of the pattern speeds of bars indicate that nearly all galaxies have
% $1 < \Rot < 1.4$\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
% 2019MNRAS.482.1733G, 2020MNRAS.491.3655G}.

% The role of gas on the evolution of the bar is less well-understood. Since the
% gas phase typically only contributes about $10-20\%$ of the mass of a galaxy at
% the present day, one might naively expect it to have a subdominant effect on the
% bar. However, because gas is collisional, it can participate in non-resonant
% angular momentum exchange with the bar\cite{2011MNRAS.415.1027H}. Thus,
% numerical work has shown that the gas phase can have a stronger influence on a
% bar than its contribution to the mass of a galaxy would
% suggest\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}.

\section{Methods}
\label{sec:methods}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
    \caption{The \textit{upper panel} shows the evolution of the pattern speed.
    As expected, the bar in the \Nbody{} run slows down due to interactions
    between the bar and the dark matter halo. However, the bar in the SMUGGLE
    run does not slow down and instead remains at a constant pattern speed. The
    \textit{lower panel} shows the torque on the bar by different components.
    The solid lines indicate the torque exerted by the halo in both the \Nbody{}
    and SMUGGLE cases. The dashed line is the torque exerted by the gas phase in
    the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
    properties are calculated is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/Rb_A2.pdf}
    \caption{The evolution in bar length and strength. The \textit{upper panel}
    shows the evolution of the bar length. In the \Nbody{} case, the bar
    lengthens. This occurs because as the pattern speed drops, bar-like orbits
    at larger radii are possible. Stars are captured on these orbits,
    lengthening the bar. This process does not occur in the SMUGGLE cases since
    the bar pattern speed is not decreasing, and therefore the bar length
    remains constant. The bar strength, shown in the \textit{lower panel} is
    measured as the maximum of the second Fourier component divided by the
    zeroth Fourier component. We see that in the \Nbody{} case (blue) the bar
    strength increases with time, consistent with previous results showing that
    the bar strength increases as bars slow down. In the SMUGGLE case (orange),
    we see that the bar strength remains roughly constant, possibly slightly
    decreasing with time. This is also consistent with the expected relation
    between pattern speed and strength since the bar in this case is not slowing
    down.}
    \label{fig:strength}
\end{figure}
\subsection{Initial Conditions}
The initial setup of the galactic disk used in this work follows closely the
GALAKOS model \citep{2020ApJ...890..117D}, which uses a modified version of the
\texttt{MakeNewDisk} code \citep{2005MNRAS.361..776S}. The GALAKOS model has
three components - a radially exponential and vertically isothermal stellar
disk, and a stellar bulge and dark matter halo following a Hernquist profile
\citep{1990ApJ...356..359H}. All \Nbody{} runs in this work used the same setup
parameters as the GALAKOS disk, more details of which can be found in the
original paper.

The addition of the gas phase was done as follows. The version of
\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas disk
which is radially exponential and in vertical gravito-hydrodynamic balance. We
modified the radial profile of this code in order to allow us to generate a disk
with a constant surface density within some cut-off radius, and then
exponentially declining beyond that radius with the scale-length of the stellar
disk. Our fiducial model used an initial surface density of
$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\textrm{kpc}$. The
initial gas disk is generated with a temperature of $10^4\,\textrm{K}$ and solar
metallicity.

After generating the gaseous disk in this way, we stitched the gas disk together
with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after the
GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The purpose of
allowing the GALAKOS disk to evolve first for a short period of time is to allow
for the bar to form unimpacted by the presence of the gas. We found that
including the gas before the bar has formed disrupts the formation of the bar,
as has been seen previously \citep[e.g.,][]{2013MNRAS.429.1949A}. Throughout
this work, we consider $t=0$ for the \Nbody{} run to be the time at which we
added the gas phase for the SMUGGLE run (i.e., we ignore the first
$1.5\,\textrm{Gyr}$ of evolution of the \Nbody{} disk when the bar is forming).

We made one additional modification when stitching the gas disk together with
the \Nbody{} disk - we created a hole within the central $4\,\textrm{kpc}$ of
the gas disks. This hole guards against an initial dramatic infall of gas within
the bar region, which we found to destroy the bar. It is not uncommon for
observed barred galaxies to have gas deficits in the bar region \citep[though
not in the very center;][]{1993RPPh...56..173S}. Therefore, our practice of
allowing the gas distribution to have a hole in the central region is consistent
with our choice to begin the simulations with a bar already formed. In this
manner, we are able to study the ensuing self-consistent interaction between the
bar and the gas, but of course we are unable to explore the origin of bars in
the presence of the gas.

We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
components (initial stellar disk, stellar bulge, and gas) and a mass resolution
of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass resolution is
closest to ``level 3'' in the AURIGA simulations \citep{2017MNRAS.467..179G}.
This corresponds to approximately $6.4\times10^6$ particles in the stellar disk,
$1.1\times10^6$ in the bulge, $1.2\times10^6$ in the gas disk, and
$25.3\times10^6$ in the dark matter halo. We used a softening length of
$0.02\,\textrm{kpc}$ for all components. Snapshots were saved at equal intervals
of $0.005$ in the time units of the simulation,
$\textrm{kpc}/(\textrm{km}/\textrm{s})$.

Our setup is initially out of equilibrium, but we found that after about
$500\,\textrm{Myr}$, the system has settled into a roughly steady-state
configuration and initial transients appear not to affect the results after this
point. The constant surface density of the initial gas disk is important for
ensuring that the gas disk is dense enough in order for comparisons to real
galaxies to be appropriate.

\subsection{Numerical Model}
We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
\citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
model additionally includes radiative heating and cooling, star formation, and
stellar feedback. Explicit gas cooling and heating of the multi-phase
interstellar medium is implemented, covering temperature ranges between $10$ and
$10^8\,\textrm{K}$.

Star formation occurs in cells above a density threshold
($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
$\epsilon = 0.01$. Star formation converts gas cells into star particles which
represent single stellar populations with a Chabrier initial mass function
\citep{2003PASP..115..763C}. For each star particle, the deposition of energy,
momentum, mass, and metals from stellar winds and supernovae is modeled.
Photo-ionization and radiation pressure are handled using an approximate
treatment. A more detailed description of this model can be found in the
flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.

We used the fiducial model parameters, except that we increased the number of
effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from $64$
to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
inefficient photo-ionization feedback since the photo-ionizing budget had not
been exhausted after deposition into $64$ neighboring cells. We also employed an
updated version of SMUGGLE using a new mechanical feedback routine similar to
the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
tensor renormalization which ensures linear and angular momentum conservation to
machine precision.

In addition to the SMUGGLE model, we considered a simpler model of the
interstellar medium based upon \citet{2003MNRAS.339..289S}. In this approach,
the multiphase nature of the interstellar medium is described in a subgrid
manner by allowing each resolution element to have a ``cold'' and ``hot''
component, with the equation of state of the gas suitably modified. Gas is
allowed to interchange between the cold and hot components through processes
such as cooling and stellar feedback. Cold gas is allowed to undergo star
formation. We refer to this model as the smooth interstellar medium model, and
it is described in more detail in \citet{2019MNRAS.489.4233M}.

\subsection{Bar Analysis}
\label{ssec:bar_analysis}
The analysis of various bar properties is performed as follows. First, the
pattern speed is measured from the angle of the second Fourier component. We
measured the second Fourier component by computing,
\begin{equation}
\begin{split}
A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
A_0 &= \sum_i m_i \textrm{,}
\end{split}
\end{equation}
where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
After correcting for the periodicity of $\phi_b$, we measured the pattern
speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
time.

In order to compute other properties of the bar, it is necessary to decompose
the disk into a barred and unbarred component. We achieved this by following
closely the methods described in \citet{2016MNRAS.463.1952P}. Our implementation
is described in more detail in Appendix~\ref{app:bardecomp}. After the disk has
been decomposed into a trapped and untrapped component, we measured the bar
length as being the radius $R_b$ which encapsulates $99\%$ of the stars
identified as being trapped in the bar.

To compute torques we used the tree algorithm in \texttt{MakeNewDisk}
\citep{2005MNRAS.361..776S} customized to be accessible from \texttt{Python}
using \texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code
\citep{1989ApJS...70..419H}. We constructed a tree with an opening angle of
$0.35$ using only the star particles identified as being trapped in the bar. We
then queried the tree at the locations of all resolution elements in the other
components and computed the torque of the bar on such components. The torque on
the bar by the other components is simply the negative of the torque on the
other components by the bar.

\subsection{Plotting Details}
We saved snapshots in intervals of $0.005$ in the time units of the
simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
throughout this work we referred to the native code time unit as
$\textrm{Gyr}$. None of our results are sensitive to this choice.

In order to remove numerical noise in several quantities we computed, we
applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
filter was applied to plots of pattern speeds, bar lengths and strengths,
torques, and angle differences.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/halo_wake.pdf}
    \caption{The wake excited in the dark matter halo. The dark matter halo wake
    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
    \textit{middle} and \textit{right panels} show a surface density projection
    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
    has been subtracted. The solid line indicates the direction of the bar while
    the dashed line indicates the direction of the halo wake (both measured by
    taking the second Fourier component within a sphere of all material within a
    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
    evolution of the angle difference between the bar and the halo wake, as
    measured from the second Fourier component. After the first Gyr, the angle
    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
    \label{fig:wake}
\end{figure*}

\section{Results}
\label{sec:results}

We present the time evolution of different bar properties in Fig.~\ref{fig:prop}.
In the upper panel, we show the pattern speed over time in the \Nbody{} (blue)
and SMUGGLE (orange) runs. The pattern speed in the \Nbody{} case slows down
while the pattern speed in the SMUGGLE case remains roughly constant. The
slowing down of the pattern speed in the \Nbody{} case is consistent with a long
line of numerical research on bars in \Nbody{} simulations
\citep{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
2009ApJ...697..293D}.

However, in the SMUGGLE case the pattern speed remains constant. After the first
Gyr of evolution, we find that the pattern speed increases by only $\sim10\%$
over the next $4\,\textrm{Gyr}$, compared to a $\sim43\%$ decrease in the
pattern speed for the \Nbody{} run over the same interval.

The bottom panel of Fig.~\ref{fig:prop} shows the torque exerted on the bar by
different components. The solid lines indicate the torque on the bar by the dark
matter halo, whereas the dashed line is the torque on the bar by the gas phase.
In the \Nbody{} case, the halo exerts a steady negative torque on the bar, with
an average torque from $1$ to $4\,\textrm{Gyr}$ of $-58.0$ in units of
$10^{10}M_{\odot}\,(\textrm{km}/\textrm{s})^2$. The halo in the SMUGGLE case
exerts a similar negative torque on the bar in the first Gyr of evolution, but
after that the halo exerts a much weaker torque on the bar, averaging only
$-7.8$ in the same units and over the same time interval. The gas in the SMUGGLE
case exerts a steady positive torque averaging $11.7$ over $1\,\textrm{Gyr}$ in
the same units.

As we saw qualitatively in Fig.~\ref{fig:overview}, the upper panel of
Fig.~\ref{fig:strength} shows that the length of the bar in the \Nbody{} case
grows over time while it remains roughly constant in the SMUGGLE case. This is
also consistent with previous numerical work, which found that bars tend to grow
as they slow down and the radius of corotation increases
\citep{2000ApJ...543..704D, 2003MNRAS.341.1179A}. The time evolution of the bar
strength, defined as the maximum of $\left|A_2/A_0\right|$ as a function of
radius, is shown in the lower panel of Fig.~\ref{fig:strength}. The quantity
$\left|A_2/A_0\right|$ varies from $0$ to $1$, with larger values indicating a
stronger bar pattern. We see that in the \Nbody{} case, $\left|A_2/A_0\right|$
increases over time as the bar pattern slows. This is consistent with previous
\Nbody{} simulations which showed a clear correlation between the bar pattern
speed and the bar strength \citep[e.g.,][]{2003MNRAS.341.1179A}. In the SMUGGLE
case, we see that the bar strength has an initial drop but then remains at a
roughly constant, but slightly decreasing, strength. This is consistent with the
pattern speed in the SMUGGLE case being roughly constant or slightly increasing.

\section{Discussion}
\label{sec:discussion}
\subsection{Pattern Speed Evolution}
The lack of evolution in the pattern speed of the SMUGGLE case (seen in
Fig.~\ref{fig:prop}) is intimately tied to the sudden decrease in torque exerted
on the bar by the dark matter halo. We argue that this can be understood in
terms of the halo wake mechanism. In the \Nbody{} case, halo material that is
resonant with the bar will form a wake which lags behind and exerts a negative
torque on the bar, slowing it down \citep{1984MNRAS.209..729T,
1985MNRAS.213..451W, 1992ApJ...400...80H}.\footnote{Since the bar is not a solid
body, it is not guaranteed that a negative torque will slow it down - e.g. a
negative torque could shred the bar, reducing its moment of inertia without
changing its pattern speed. However, the bar seems to empirically respond to a
negative torque induced by a halo wake by slowing down.} As the bar slows down,
the location of the resonances in phase space changes, allowing halo material
newly resonant with the bar to participate in forming the wake. However, the gas
is a reliable source of positive torque on the bar, speeding the bar up. This
stops the resonance location from changing such that the halo cannot reinforce
the wake, arresting the process by which the halo can slow the bar down. We term
this process ``halo exhaustion.''

We can test this interpretation by measuring the angle offset between the halo
wake and the bar. If the wake and the bar are aligned (i.e., there is no angle
offset), then the wake cannot exert a negative torque on the bar. This angle
is plotted in the left panel of Fig.~\ref{fig:wake}, which shows that the
angle offset is larger in the \Nbody{} case than in the SMUGGLE case by about a
factor of two. The center and right panels of Fig.~\ref{fig:wake} show the halo
wake with respect to the location of the bar in the \Nbody{} (center) and SMUGGLE
(right) cases at one point in time.

The presence of the gas can arrest the process by which additional material in
the dark matter halo can contribute to a wake. However, this does not explain
why the pattern speed in the SMUGGLE case is nearly constant over several Gyr.
Naively, it would be a coincidence that the bar pattern speed remains constant
in the SMUGGLE case, resulting from a chance cancellation of the halo and gas
torques. However, a constant pattern speed in the presence of gas has been
observed in a few simulations of barred galaxies with
gas \citep{1993AA...268...65F, 2010ApJ...719.1470V, 2014MNRAS.438L..81A}. Previous
work has argued this is due to the bar torquing gas inwards, but no explanation
has been given for why it might remain constant.

We propose that an equilibrium mechanism is responsible for the pattern speed
remaining approximately constant. In this scenario, residual negative torque
from the dark matter halo balances out the positive torque from the gas phase.
It has been shown when an analytic bar is forced to rotate at a constant pattern
speed for a few Gyr, the halo exerts almost no torque on the bar
\citep{2022MNRAS.513..768C}. We saw in Fig.~\ref{fig:prop} that the dark matter
halo in our simulation is still able to support some negative torque over a
several Gyr time span.

We argue that the following occurs. First, the bar is not able to slow down
quickly enough due to the positive torque of the infalling gas. This causes
the resonant halo phase space at a particular pattern speed, $\Omega_{\textrm{p},0}$,
to become exhausted and no longer able to support a negative torque. Second,
the gas is still exerting a positive torque on the bar, and therefore the
pattern speed will again increase. Since at higher pattern speeds the halo has
not yet been totally exhausted, the halo will once again be able to exert a
negative torque on the bar. The pattern speed will then settle at a new value
slightly higher than $\Omega_{\textrm{p},0}$ where the gas and halo torques cancel.
Over time, the pattern speed should slowly increase as the halo becomes
progressively exhausted.

\subsection{Delayed Gas Injection}
A clear prediction of our proposed mechanism is that the constant pattern speed
a particular galaxy will end up is somewhat arbitrary. In the real universe for
an isolated galaxy it would be the formation pattern speed of the bar while in
our simulation it is the pattern speed of the bar when gas is added to the
system. We tested this by adding gas to the system at a later time when the bar
has further grown and slowed down with time. In our particular test, we added
the gas at a time when the pattern speed is
$\sim30\,\textrm{km}/\textrm{s}/\textrm{kpc}$. As shown in
Fig.~\ref{fig:snap700}, we find that the pattern speed evolution is very similar
between the two cases (orange and red lines). If anything, the system with a
lower pattern speed seems to speed up more, which is consistent with our picture
since the stronger bar should experience a larger torque from the gas as it is
more efficient at driving gas inflows. We also show in the
Appendix~\ref{app:varyps} that more slowly rotating bars at fixed bar strength
are more efficient at driving gas inflows as well. Nonetheless, when the initial
pattern speed is lower (red line), the addition of gas does not cause the
pattern speed to quickly return to the higher value of our fiducial simulation
(orange line).

\subsection{Varying Initial Gas Fractions}
We performed a test in which we varied the initial gas fraction of the
disk. In our fiducial run, we set the surface density of the gas disk from
$4\,\textrm{kpc}$ to $\sim9.3\,\textrm{kpc}$ to be $20\,\Msun/\textrm{pc}^2$.
We also ran with surface densities of $15$, $10$, and
$5\,\Msun/\textrm{pc}^2$. These correspond to initial gas fractions of
approximately $16\%$, $10\%$, $7\%$, and $4\%$. The pattern speed evolution is
shown in Fig.~\ref{fig:fgas} We find that the bar in disks with
initial surface densities of $20$, $15$, and $10\,\Msun/\textrm{pc}^2$ evolve
with a constant pattern speed while the bar in a disk with initial surface
density of $5\,\Msun/\textrm{kpc}^2$ slows down at a similar rate to the \Nbody{} case.

As a result, we conclude that for the disk, bar, and halo properties
considered in this work, a gas fraction of only approximately $5\%$ is
necessary in order for the proposed stabilizing mechanism to operate.

\subsection{Smooth Interstellar Medium}
We performed a simulation of the same disk but with a simpler model of the
interstellar medium \citep{2003MNRAS.339..289S}, closer to standard methods used
in cosmological simulations of galaxy formation and described in more detail in
Section~\ref{sec:methods}. The result of this test is presented in
Fig.~\ref{fig:GFM}. We find that the pattern speed evolution is nearly the same
in this case, and so conclude that our result is not sensitive to the details of
the model for the interstellar medium.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_late_start.pdf}
    \caption{Pattern speed evolution with a lower initial pattern speed. We
    tested the evolution of our system when gas is added to the \Nbody{} run at
    a later time, but with all other simulation parameters kept the same. The
    setup is therefore identical to our previous runs just with a bar that is
    larger, stronger, and with a lower pattern speed. We find that the pattern
    speed evolution is very similar to our fiducial case, except that the bar
    retains its original pattern speed. This indicates a mechanism which keeps
    the bar at its formation pattern speed, and that there is not a particular
    pattern speed which the system tends to.}
    \label{fig:snap700}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_fgas.pdf}
    \caption{Pattern speed evolution with varying gas fractions. We explored the
    impact of lowering the initial gas surface density of our fiducial disk on
    the evolution of the pattern speed. The surface densities we tested of $20$,
    $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to initial gas
    fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that initial surface
    densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in bars which
    remain at a constant pattern speed, while an initial surface density of
    $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in roughly the
    same manner as the \Nbody{} case.}
    \label{fig:fgas}
\end{figure}

\subsection{Semi-Analytic Model}
We also developed a simple semi-analytic model of a bar-disk-halo system. This
exercise demonstrates that our proposed mechanism follows from a few simple
assumptions. Our method follows closely the one developed in
\citet{2022MNRAS.513..768C}. We model the bar-disk-halo system with three
components: a dark matter \citet{1990ApJ...356..359H} halo, a
\citet{1975PASJ...27..533M} disk, and a pure quadrupole bar
described in \citet{2022MNRAS.513..768C}. The bar and disk components are just
described by their potential, but we integrate the trajectories of test
particles drawn from a Hernquist halo. Note that we do not include the
interactions between these test particles, and so this model is not
self-consistent. We give our chosen parameter values in Appendix~\ref{app:sam}.

We allow the bar in this model to rotate as a solid body. However, we crucially
allow the pattern speed of the bar to freely change with time in accordance with
the torque exerted on the dark matter halo by the bar. In particular, we
subtract the $z$-component of this torque divided by the moment of inertia of
the bar from the pattern speed at each timestep. Since the radius of corotation
\RCR{} is a parameter in the bar model from \citet{2022MNRAS.513..768C}, we
allow the moment of inertia of the bar to vary with ${\RCR}^2$. To be more
precise, we allow
\begin{equation}
I = \frac{I_6}{10^{10} \Msun (\kms)^2} \left( \frac{\RCR}{6\,\textrm{kpc}} \right)^2\textrm{,}
\end{equation}
where $I_6$ is a free parameter chosen by the user. We found that allowing
$I_6=8$ is a good approximation to our fiducial disk model. In code units, the
moment of inertia of the SMUGGLE bar (i.e., the particles classified as being in
the bar) is about $2$. This is a factor of $4$ smaller than our fiducial value
of $I_6=8$, but this is probably due either to the fact that the bar does not
really rotate as a solid body or that resonantly captured stars contribute to
the real bar's effective moment of inertia \citep{1985MNRAS.213..451W}.

In addition to $I_6$, we allowed for another free parameter - the torque from
the gas phase on the bar, $\tau_{\textrm{gas}}$. This torque is applied to the
bar in the same way as the torque from the halo is applied. The torque is given
in code units ($10^{10}\Msun \left(\kms\right)^2$). We allow this torque to
mildly vary with pattern speed according to the formula,
\begin{equation}
\tau = \tau_{\textrm{gas}} + (40 - \Omega_{\textrm{p}})\textrm{,}
\end{equation}
where $\Omega_{\textrm{p}}$ is given in $\kms/\textrm{kpc}$. We are motivated to
do this because of our result in Appendix~\ref{app:varyps} which shows that more
slowly rotating bars have higher gas infall rates and thus larger positive
torques. In the $\tau_{\textrm{gas}}=0$ case, we do not allow the gas torque to
vary with pattern speed.

We show the effect of varying the gas torque $\tau_{\textrm{gas}}$ from $0$ to
$20$ in increments of $2$ in Fig.~\ref{fig:sam}. The solid lines indicate the
semi-analytic model, while the two dashed lines correspond to our fiducial
simulations introduced earlier. For reference, the average torque exerted by the
gas phase on the bar in our fiducial simulation was $11.7$ in code units. We see
in Fig.~\ref{fig:sam} that we can reproduce the stability of our fiducial gas
disk (i.e., its lack of secular evolution) simply by including a positive torque
on the order of $6$. Our semi-analytic model with no gas torque can reproduce
the pattern speed evolution of the $N$-body case.

We next take the $\tau_{\textrm{gas}}=0$ and $20$ cases from Fig.~\ref{fig:sam}
and plot the halo torque evolution. This result, given in Fig.~\ref{fig:sam-torque},
is comparable to the lower panel of Fig.~\ref{fig:prop}. We find that the
$\tau_{\textrm{gas}}=0$ case compares favorably to the \Nbody{} case described
in previous sections. The bar exerts a steady negative torque in this case (blue
line). When a gas torque is included (orange lines), we find that the halo's
torque becomes much weaker, similar to what we found in the SMUGGLE case. The
gas applies a steady positive torque by construction. Therefore, the exhaustion
process is reproduced in our simple semi-analytic model.

\subsection{Observations}
Observational estimates of the pattern speeds of bars indicate that nearly all
galaxies have $1 < \Rot < 1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A,
2019MNRAS.482.1733G, 2020MNRAS.491.3655G}, where $\Rot$ was defined in
Section~\ref{sec:intro} to be $\Rot\equiv \RCR/\Rb$. This observational fact has
long been in conflict with the theoretical expectation that bars should slow
down, increasing \Rot{} \citep[e.g.][]{1984MNRAS.209..729T, 1985MNRAS.213..451W,
2000ApJ...543..704D}. Explanations for this discrepancy have been given in the
past. Some have argued that perhaps the central regions of dark matter haloes
are less dense than we expected from $\Lambda\textrm{CDM}$
\citep[e.g.][]{2000ApJ...543..704D,2021AA...650L..16F}. Some have argued that
perhaps bars are recurrent, short-lived phenomena, and that all the bars we see
in the local universe are very young \citep{2002AA...392...83B,
2005MNRAS.364L..18B}. Some have argued that modifications to General Relativity
ease the tension between the observed universe and $\Lambda\textrm{CDM}$
\citep[e.g.][]{2021MNRAS.503.2833R, 2021MNRAS.508..926R}.

Because such a small gas fraction is necessary for our stabilizing mechanism to
operate ($5\%$ in our Milky Way-like disk), we argue that most galaxies host a
bar that is not slowing down. This naturally explains why most observed bars are
fast rotators. However, we acknowledge two instances of reported discrepancies
between our mechanism and observations.

First, we note that \citet{2020MNRAS.491.3655G} found that the rotation
parameter \Rot{} positively correlates with gas fraction, such that galaxies
with higher gas fractions are rotating more slowly. However, it is not obvious
this is in tension with our result since the gas fraction of galaxies correlates
with other galactic properties \citep{2009ARAA..47..159B}. Furthermore, the
measurement of pattern speeds is a delicate process still prone to large errors.

Second, the work of \citet{2021MNRAS.500.4710C} and \citet{2021MNRAS.505.2412C}
have made indirect measurements of the deceleration of the bar's pattern speed
from kinematics and chemistry. We point out that these reported measurements are
not direct measurements of the Milky Way bar's deceleration. Much like the
simulations in the present work, the simulations of these two works do not
properly account for the complicated formation process of the Galactic bar,
which may leave imprints on the present day distribution of stars in spatial,
kinematic, and chemical space. More investigation is necessary to reconcile the
present work with these two well-executed manuscripts.

\subsection{Cosmological Simulations}
Barred galaxies in cosmological simulations of galaxy formation continue to be
in conflict with observations by producing bars which rotate too slowly
\citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F,
2022ApJ...940...61F}. However, the pattern speeds of bars in both cosmological
simulations and the real universe can be affected by environmental processes not
included in our simulation -- e.g., satellite infall
\citep{2011Natur.477..301P}, non-sphericity \citep{2013MNRAS.429.1949A},
rotation in the dark matter halo \citep{2013MNRAS.434.1287S,
2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C}, or perhaps even
the gaseous circumgalactic medium. Naturally, extending our present work to
account for such effects is a crucial next step in understanding the formation
and evolution of galactic bars. We are presently engaged in such an exploration.
We note that \citet{2022ApJ...940...61F} argue that bars in the TNG50 simulation
are too short rather than too slow.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/ps_GFM.pdf}
    \caption{Pattern speed evolution of a smooth ISM model. This evolution is
    shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE (orange), and
    smooth ISM (red) cases. The smooth ISM model is an older model for the ISM
    which treats its multiphase nature in a subgrid fashion
    \citep{2003MNRAS.339..289S}. This fundamentally differs from the SMUGGLE
    model, which explicitly resolves the hot and cold phases of the ISM
    \citep{2019MNRAS.489.4233M}. The pattern speed in the smooth ISM case is
    broadly similar to the evolution in the SMUGGLE case. This shows that the
    stability of the pattern speed is not simply a result of our assumed model
    for the ISM.}
\label{fig:GFM}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/samGvar.pdf}
    \caption{A comparison between the pattern speeds of our fiducial disk
    systems and a semi-analytic model. The solid lines indicate the pattern
    speeds assuming a constant positive torque, varying in increments of $2$
    from $0$ to $20$. The dashed lines indicate the pattern speed evolution from
    our fully self-consistent simulations from earlier. We find excellent
    agreement between our fiducial simulations and our semi-analytic model of a
    bar-disk-halo system. Torques are given in code units ($10^{10}\Msun
    \left(\kms\right)^2$).}
    \label{fig:sam}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/sam_torque.pdf}
    \caption{The torque exerted on the bar by various components in our
    semi-analytic model. Solid lines indicate the torque by the halo while the
    dashed line indicates the torque exerted by the gas phase. We chose two
    models with $\tau_{\textrm{gas}}=0$ and $20$, which most closely resemble
    our $N$-body and SMUGGLE disks, respectively. This figure ought to be
    compared to the lower panel of Fig.~\ref{fig:prop}. Overall, we find good
    qualitative agreement.}
    \label{fig:sam-torque}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
We performed a simulation of a Milky Way-like galactic disk hosting a strong bar
with a state-of-the-art model for the interstellar medium. We found that the
pattern speed of the bar in this simulation does not slow down but rather
remains at a stable, constant pattern speed. We provided a simple semi-analytic
model which reproduces many of the features from our fiducial disk model.

The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense \citep{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics \citep{2021MNRAS.503.2833R,
2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar \citep{2021MNRAS.500.4710C,
2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time Our work does not alter
expectations for radial mixing induced by spiral arms
\citep{2002MNRAS.336..785S}.

We found that below a certain gas fraction, bars should still be able to slow
down. Therefore, we expect barred spiral galaxies which have been gas-poor for
extended periods of time to be rotating very slowly. We therefore predict that
observations which target such galaxies (e.g., lenticular barred galaxies
\citep{2009ARAA..47..159B}) would find slowly rotating bars.\footnote{This does
not mean that we predict \textit{all} gas-poor galaxies should be slowly
rotating. Indeed, they would need to be gas-deficient for several $\textrm{Gyr}$
before they would be classified as slow rotators.} There does exist one example
of a galaxy which is known to be a slow rotator -- the low surface brightness
galaxy UGC 628 \citep{2009AA...499L..25C}. This galaxy has been studied in
detail by \citet{2016MNRAS.463.1751C}, who note that it indeed has a low gas
fraction for galaxies of its type. We predict a general trend that bars in
gas-rich spiral galaxies should rotate quickly while some bars in gas-poor
spiral galaxies should rotate slowly.

\begin{acknowledgments}
  We would like to thank Greg~L. Bryan, Neal~J. Evans, Drummond~B. Fielding, Keith
  Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V. Johnston,
  Peter~M.~W. Kalberla, Jürgen Kerp, Julio~F. Navarro, Joshua~S. Speagle,
  Martin~D. Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to
  thank Todd Phillips for helpful discussions. Some computations in this paper
  were run on the FASRC Cannon cluster supported by the FAS Division of Science
  Research Computing Group at Harvard University. Resources supporting this work
  were also provided by the NASA High-End Computing (HEC) Program through the NASA
  Advanced Supercomputing (NAS) Division at Ames Research Center. This research
  has made use of the Spanish Virtual Observatory (https://svo.cab.inta-csic.es)
  project funded by MCIN/AEI/10.13039/501100011033/ through grant
  PID2020-112949GB-I00. A.B. was supported by the Future Investigators in NASA
  Earth and Space Science and Technology (FINESST) award number 80NSSC20K1536
  during the completion of this work. E.D. was partially supported by HST grants
  HST-AR-16363.001 and HST-AR-16602.006-A and by NASA Award NASA 80NSSC22K0761.
  J.Q. acknowledges support from NSF grant AST-2008490. L.V.S. is grateful for
  financial support from NASA ATP 80NSSC20K0566, NSF AST 1817233 and NSF CAREER
  1945310 grants. P.T. acknowledges support from NSF grant AST-1909933,
  AST-2008490, and NASA ATP Grant 80NSSC20K0502. M.V. acknowledges support through
  NASA ATP 19-ATP19-0019, 19-ATP19-0020, 19-ATP19-0167, and NSF grants
  AST-1814053, AST-1814259, AST-1909831, AST-2007355 and AST-2107724.
\end{acknowledgments}

\software{
{\sc agama} \url{https://github.com/GalacticDynamics-Oxford/Agama}, {\sc
astropy} \citep{astropy:2013, astropy:2018}, {\sc h5py}
\url{http://www.h5py.org/}, {\sc inspector\_gadget}
\url{https://bitbucket.org/abauer/inspector_gadget/}, {\sc joblib}
\url{https://joblib.readthedocs.io/en/latest/}, {\sc matplotlib}
\citep{Hunter:2007}, {\sc numba} \citep{Lam:2015bfh}, {\sc numpy}
\citep{harris2020array}, {\sc scipy} \citep{2020SciPy-NMeth}, {\sc tqdm}
\url{https://tqdm.github.io/}
          }

\appendix

\section{Bar Decomposition}
\label{app:bardecomp}
\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/bar_decomp.pdf}
    \caption{Disk decomposition into the barred and unbarred disk. This
    procedure is based on \citet{2016MNRAS.463.1952P}. The \textit{left panel}
    shows a face-on surface density projection through the stellar component of
    the \SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
    \textit{middle panel} shows the component of the disk identified as being
    trapped in the bar while the \textit{right panel} shows the component of the
    disk identified as not being trapped in the bar. The fact that the untrapped
    stars form a roughly axisymmetric structure indicates our bar decomposition
    is sufficiently accurate. We have computed that $76\%$ of the second Fourier
    component resides in the stars classified as being trapped in the bar.}
    \label{fig:decomp}
\end{figure*}

Computing the length of the bar and the torque on the bar by different
components requires us to decompose the disk into a component which is trapped
by the bar and a component which is untrapped. In order to do this, we follow
closely the technique developed in \citet{2016MNRAS.463.1952P}. We analyzed the
orbit of each star particle (meaning initial disk, bulge, and newly formed
stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
corotating with the bar, where apoapses are defined as local maxima in $r$. For
each apoapse, we searched for the $19$ closest apoapses in time and applied a
$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
computed for each of the two clusters the average angle from the bar
$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
${\sigma_R}_{0,1}$, and the average radius of the cluster
$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in the
bar if it met the following criteria:
\begin{equation}
\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
\end{equation}
\begin{equation}
\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
\end{equation}
These criterion are slightly different and simplified from the ones used in
\citet{2016MNRAS.463.1952P}, but we found them to empirically work well at
decomposing the disk into a bar and disk component. In Fig.~\ref{fig:decomp}, we
show an example of this decomposition. The \textit{left} panel shows a surface
density projection of the stellar disk and bulge (including newly formed stars)
from the SMUGGLE model after $1\,\text{Gyr}$ of evolution in a frame such that
the bar is aligned with the $x$-axis. The \textit{middle} panel shows a
projection of the subset of stars that are identified as being trapped in the
bar and the \textit{right} panel shows a projection of the stars that are not
identified as being trapped. The fact that the \textit{right} panel is roughly
axisymmetric indicates the bar decomposition is performing adequately.

We computed the second Fourier component $A_2$ for all particles classified as
barred and unbarred. We found that $76\%$ of the total $m=2$ Fourier component
is in the particles classified as barred (i.e.,
$A_{2,\textrm{bar}}/A_{2,\textrm{tot}}\sim0.76$). This indicates good but not
perfect classification. Some of this is probably coming from the $m=4$ component
being classified as unbarred, given that the unbarred component has a somewhat
boxy shape.

\section{Varying Pattern Speed}
\label{app:varyps}
When the bar slows down, we argue that this induces a larger positive torque
from the gas phase. Only gas within corotation will flow inwards, while gas
outside corotation will flow outwards \citep{2011MNRAS.415.1027H}. Since the
corotation radius is larger for more slowly rotating bars, it follows that
more slowly rotating bars should be more efficient at driving gas inflows and
thus experience a larger positive torque from the gas phase.

We performed an experiment to test this hypothesis by forcing the stellar disk
in the SMUGGLE run to rotate at a constant angular rate and measuring the
torque on the bar by the gas phase at different rotation rates. The result of
this experiment is illustrated in Fig.~\ref{fig:equil}, which shows that a more
slowly rotating bar experiences a larger positive torque from the gas.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/torque_ps.pdf}
    \caption{Average torque exerted by gas on a bar which rotates at a fixed
    pattern speed. Since only gas within the corotation radius is able to infall
    and slower bars have larger corotation radii, slower bars experience a
    larger net torque than faster bars. The setup of the simulations used here
    is identical to the SMUGGLE case discussed earlier, except the \Nbody{} disk
    is rotated as a solid body with a constant angular
    velocity.}
    \label{fig:equil}
\end{figure}

\section{Semi-Analytic Model Parameters}
\label{app:sam}
Our semi-analytic model consisted of a three-component bar-disk-halo system. We
describe here the parameters we chose for these components. The parameters of
the disk and halo were chosen to match closely what we used in our fiducial
simulations. The system can thus be understood as being roughly similar to the
Milky Way, though no careful analysis has been performed to ensure the closest
match possible.

For the dark matter halo, we used a Hernquist potential
\citep{1990ApJ...356..359H} with mass $10^{12}\,\Msun$ and a scale length of
$26.2\,\textrm{kpc}$. For the stellar disk, we used a MiyaMoto-Nagai disk
\citep{1975PASJ...27..533M} with mass $4.8\times10^{10}\,\Msun$, radial scale
length of $2.67\,\textrm{kpc}$, and vertical scale length of
$0.32\,\textrm{kpc}$. For the bar, we used the quadrupole potential described in
\citet{2022MNRAS.513..768C}. We used their fiducial parameter values --
specifically, we set $A=0.02$, $b=0.28$, and $v_c = 235\,\kms$. Our initial
pattern speed is always set to $40\,\kms/\textrm{kpc}$.

We integrated our model for $5\,\textrm{Gyr}$ with a timestep of
$0.01\,\textrm{Gyr}$.

\section{Idealized Mock Light Images}
\label{app:hst}
We created synthetic mock light images at $1024^2$ resolution in
Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
including secondary emission from dust \citep{2020AC....3100381C}. Emission from
star particles was calculated assuming a Bruzual-Charlot spectral energy
distribution \citep{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
function \citep{2003PASP..115..763C}. The luminosity from each star particle was
smoothed using a cubic spline kernel with each star particle being assigned a
smoothing length equal to its distance to its $32$nd closest neighbor. We
launched $10^9$ photon packets per simulation segment. For star particles
which were present in the initial simulation (i.e., initial disk and bulge
particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
metallicity. For star particles in the SMUGGLE simulation which are newly
formed, we used their recorded ages and metallicities.

For the SMUGGLE simulations, we included absorption and emission from dust. We
assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
Solids) dust model \citep{2017AA...602A..46J} to convert gas metallicity and
mass density to dust mass density. For the \Nbody{} simulations we ran the
SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
same procedure.

We used the profiles for filters F814W, F606W, and F475W on the Advanced
Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
scaling following \citet{2004PASP..116..133L} Scalings and cutoff values
were adjusted manually to create visually appealing images. Filter profiles
were provided by the Spanish Virtual Observatory Filter
Service \citep{2012ivoa.rept.1015R, 2020sea..confE.182R}.

\section{Comparison to the Milky Way}
\label{app:milkyway}
For several Gyr, our fiducial disk exhibits several properties in reasonable
agreement to the Milky Way. This is uncommon in models of galaxies that include
the gas phase of the disk but no circumgalactic medium. As mentioned earlier,
the pattern speed seems to match the observed pattern speed of the Milky Way's
bar \citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our
disk is comparable to the Milky Way.

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package \citep{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
bulge, gas, and newly formed stars) with an axisymmetric cylindrical spline with
$20$ grid points in both the radial and vertical direction spanning $0.2$ to
$50\,\textrm{kpc}$ in the radial direction and from $0.02$ to $10\,\textrm{kpc}$
in the vertical direction. We fit the dark matter halo using a spherically
symmetric multipole fit with a maximum angular harmonic coefficient of $l=2$. We
plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
observational estimates \citep{2019ApJ...871..120E}. The SMUGGLE disk (which
includes additional mass in the form of gas) has a slightly higher circular
velocity than the \Nbody{} disk which, itself, is slightly higher than the
observational estimates. Overall, though, the circular velocity curves between
our model and that observed in the Milky Way are broadly consistent.

We also show the evolution of the surface density profile in Fig.~\ref{fig:surf}
We find that in our simulation the atomic and molecular gas surface density and
the SFR surface density is broadly consistent with the expected values for the
Milky Way \citep{2008AA...487..951K,2022ApJ...929L..18E}. The discrepancy between
$1$ and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
to the fact that the distances to molecular clouds which underlines this work
used a simple kinematic distance based on an axisymmetric model of the Milky
Way \citep{2017ApJ...834...57M}, which is not accurate in the bar region where gas
has large non-circular velocities.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/surf_dens.pdf}
    \caption{The time evolution of the atomic gas surface density
    (\textit{upper}), molecular gas surface density (\textit{middle}) and the
    star formation rate (SFR) surface density (\textit{lower}) at various times
    during our fiducial simulation. Colored lines indicate the profiles at
    selected times during the simulation while the black dashed lines indicate
    observations for the atomic gas \citep{2008AA...487..951K} and black dotted
    lines indicate a model which allows the CO-to-H$_2$ conversion factor
    $X_{\textrm{CO}}$ to vary with metallicity \citep{2022ApJ...929L..18E}.
    Molecular gas surface densities were provided separately (N. Evans, private
    communication). We see that the molecular gas and SFR surface densities are
    within an order of magnitude of the Milky Way's typical values at all times.
    We see a sharp decrease in the gas and SFR surface densities along the
    extent of the bar from $\sim1$ to $\sim4\,\textrm{kpc}$, related to the gas
    inflow in this region.}
    \label{fig:surf}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=9cm]{fig/vcirc.pdf}
    \caption{The circular velocity curve of our initial setups. This
    curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
    compared to observational estimates for the Milky
    Way \citep{2019ApJ...871..120E}. We see that the circular velocity curve for
    both runs is marginally larger than the Milky Way's, but still comparable. The
    SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
    additional mass in the gas phase.}
    \label{fig:vcirc}
\end{figure}

%% For this sample we use BibTeX plus aasjournals.bst to generate the
%% the bibliography. The sample631.bib file was populated from ADS. To
%% get the citations to show in the compiled file do the following:
%%
%% pdflatex sample631.tex
%% bibtext sample631
%% pdflatex sample631.tex
%% pdflatex sample631.tex

\bibliography{ref}{}
\bibliographystyle{aasjournal}

%% This command is needed to show the entire author+affiliation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample631.tex'.